---
title: "Final Project Draft"
author: "Kyla Beasley, Stephanie Boehm, Krys Morawa, and Daria Petrukhina"
date: "11/24/2020"
output: html_document
---


```{r, echo = FALSE, results= FALSE}
#Set Working Directory
setwd("C:/Users/Stephanie/Documents/STAT382")
#Read data in data frame
df <- read.csv('COVID-19_Daily_Cases__Deaths__and_Hospitalizations.csv', sep = ",")
```

Check for missing data
```{r, results = FALSE}
str(df) # 5 missing values each for 19 different variables 
df[is.na(df) == TRUE] <- 0 # change NA values to 0
```

------

### Compute summary statistics 

```{r, echo = FALSE, results = FALSE}
#Function for skew
skew <- function(x){
  m3 <- sum((x-mean(x))^3)/length(x)
  m3/(sd(x)^3)
}

#Function for kurtotis
kurt <- function(x){
  m4 <- sum((x-mean(x))^4)/length(x)
  m4/(sd(x)^4)
}

``` 

```{r, results = FALSE}
summary(df)
```

```{r}
sapply(df[c(6, 10, 24, 28)], skew)
sapply(df[c(6, 10, 24, 28)], kurt)
sapply(X = df[c(6, 10, 24, 28)], hist)
sapply(df[c(6, 10, 24, 28)], boxplot)
```

None of the variables have a normal distribution because the  
skewness is not close to zero and the kurtosis is not close to  
three for any of the variables. None of the histograms look normal  
and each graph is skewed to the left and it may potentially contain  
some outlines.

------

### Hypothesis Testing

People in the age group of 18-29 are more likely to catch COVID-19 than people in the age group 60-69
People 18-29 are more likely to go out in high COVID-19 areas like bars, parties, and restaurants than
people 60-69 because people who are older tend to have other people such as other relatives like children,
grandchildren, or partners, or have issues themselves which makes it difficult to go out such as health
problems or unable to drive. People who are younger tend to have less restrictions stopping them from going out.

* $H_0:  mu_1 - mu_2 = 0$
  + The null hypothesis is that the mean number of cases for ages 18-29 is equal to those that are ages 60-69.
* $H_1:  mu_1 - mu_2 > 0$
  + The alternative is that the mean number of cases for ages 18-29 is greater than those that are ages 60-69.

```{r}
t.test(df$Cases...Age.18.29, df$Cases...Age.60.69, alternative = c("greater"))
```

The P-value is less than 0.05. Reject the null hypothesis, accept the alternative.
There is significant evidence that the mean mean number of cases for ages 18-29 is greater than those that 
are ages 60-69.

People in the age group 18-29 are less likely to die from COVID-19 than people in the age group 60-69. 
People 18-29 tend to have stronger immune systems to fight the illness than people 60-69, which causes 
there to be more deaths in the age range 60-69 than 18-29. 

* $H_0:  mu_1 - mu_2 = 0$
  + The null hypothesis is that the mean number of deaths for ages 18-29 is equal to those that are ages 60-69.
* $H_1:  mu_1 - mu_2 < 0$
  + The alternative is that the mean number of deaths for ages 18-29 is less than those that are ages 60-69.
  
```{r}
t.test(df$Deaths...Age.18.29, df$Deaths...Age.60.69, alternative = c("less"))
```

The P-value is less than 0.05. Reject the null hypothesis, accept the alternative.  
There is significant evidence that the mean mean number of deaths for ages 18-29 is less than those that 
are ages 60-69.

------

### Linear Association 

The explanatory variable is Cases Ages 18-29 and the response variable is Deaths Ages 18-29. 
The explanatory variable is Cases Ages 18-29 because a person has to have COVID in order to have that 
be the cause of death, so the response variable is Deaths Ages 18-29 since people die from COVID.

I expect there to be a moderately weak positive linear association. I expect there to be a weak 
association because if there are a lot of cases then there will be many deaths because people who are ages 
18-29 tend to have better immue systems which will help fight off the virvus and prevent death.

I expect there to be a positive association because when it is positive is means 
that the values of one variable tend to increase as the values of the other variable increase which is 
when the number of cases increase, then so do the number of deaths because there are more people likely 
to die from COVID-19 when more people have it.

I expect there to be a linear association because I expect 
there to be some type of association. 

```{r}
plot(df$Cases...Age.18.29, df$Deaths...Age.18.29) 
cor(df$Cases...Age.18.29, df$Deaths...Age.18.29)
```

The correlation coefficient is positive and close to 0, indicating a weak, positive, linear 
association between Cases Ages 18-29 and Deaths Ages 18-29.

```{r}
regline <- lm(Deaths...Age.18.29~ Cases...Age.18.29, data=df) 
summary(regline)
```

* The regression line $y= 3.752e-05x + 8.780e^-02$  
The parameter $3.752e-05$ is the slope of the equation. The parameter $8.780e-02$ is the intercept for the equation.
The X variable is the explanatory variable and the Y variable is the response variable. 

```{r}
plot(df$Cases...Age.18.29, df$Deaths...Age.18.29)
abline(regline, col="red")
```

From the plot, we can see that Deaths Ages 18-29 is constant, there was only one outlier in Deaths Ages 18-29.  
We can conclude that there is no association between Deaths Ages 18-29 and Cases Ages 18-29.  

```{r, echo=FALSE, results=FALSE}
df2 <- resid(regline) 
df2
```

```{r}
plot(df$Cases...Age.18.29, df2) 
abline(h = 0)
```
Residuals are normally distributed, and randomly scattered in the scatterplot, and they have a   
constant variance (homoscedasticity). This is consistent with what we would expect if the regression  
line were appropriate.

```{r}
hist(df2)
```
The histogram does not look normal so it is not consistent with what we would expect if the regression 
were appropriate. 

```{r}
shapiro.test(df2)
```
The p-value is less than 0.05, so reject the null hypothesis that the residuals are normal. It is not 
consistent with what we would expect if the regression were appropriate. 

--------

```{r}
model1 <- lm(Deaths...Age.18.29~Cases...Age.18.29, data = df)
summary(model1)
```


 R^2 is 0.0002392 which is close to 0 which means that the model does not fit the data well.  
 The explanatory variable is Cases Ages 60-69 and the response variable is Deaths Ages 60-69. 
 The explanatory variable is Cases Ages 60-69 because a person has to have COVID in order to have that 
 be the cause of death, so the response variable is Deaths Ages 60-69 since people die from COVID. 
 I expect there to be a moderately strong positive linear association. I expect there to be a strong 
 association because when older people tend to get sick, their immue system is not strong enough to fight off 
 the virus and causing death. I expect there to be a positive association because when it is positive is means 
 that the values of one variable tend to increase as the values of the other variable increase which is 
 when the number of cases increase, then so do the number of deaths because there are more people likely 
 to die from COVID-19 when more people have it. I expect there to be a linear association because I expect 
 there to be some type of association. 

```{r}
plot(df$Cases...Age.60.69, df$Deaths...Age.60.69) 
cor(df$Cases...Age.18.29, df$Deaths...Age.18.29)
```

 The correlation coefficient is positive and close to 0, indicating a weak, positive, linear
 association between Cases Ages 60-69 and Deaths Ages 60-69.

```{r}
regline2 <- lm(Deaths...Age.60.69~ Cases...Age.60.69, data=df) 
summary(regline2)
```

 the regression line y= 0.035688x + 1.026633
 The parameter 0.035688x is the slope of the equation. The parameter 1.026633 is the intercept for the equation.
 The X variable is the explanatory variable and the Y variable is the response variable. 

```{r}
plot(df$Cases...Age.60.69, df$Deaths...Age.60.69)
abline(regline2, col="red")
```


 from the plot, we can see that Deaths Ages 60-69 is constant, there was only one outlier in Deaths Ages 60-69
 We can conclude that there is some association between Deaths Ages 60-69 and Cases Ages 60-69

```{r, echo = FALSE, results= FALSE}
df3 <- resid(regline2) 
df3
```

```{r}
plot(df$Cases...Age.60.69, df3) 
abline(h = 0)
```

 Residuals are normally distributed, and randomly scattered in the scatterplot, and they have a  
 constant variance (homoscedasticity). This is consistent with what we would expect if the regression
 line were appropriate. 

```{r}
hist(df3)
```

 The histogram does look normal so it is consistent with what we would expect if the regression 
 were appropriate. 

```{r}
shapiro.test(df3)
```

 The p-value is less than 0.05, so reject the null hypothesis that the residuals are normal. It is not 
 consistent with what we would expect if the regression were appropriate. 

```{r}
model2 <- lm(Deaths...Age.60.69~Cases...Age.60.69, data = df)
summary(model2)
```

 R^2 is 0.206 which is closer to 0 than 1 which means that the model does not fit the data well.  

------

### ANOVA 


```{r, results = FALSE}
na.omit(df)
library(tidyverse)
dfnew <- separate(df, "Date", c("Month", "Day", "Year"), sep = "/")
dfnew
```

```{r}
boxplot(dfnew$Cases...Total~dfnew$Month)
```

 looks like there might be some differences

*  $H_0: mu_3 = mu_4 = mu_5 = mu_6 = mu_7 = mu_8 = mu_9 = mu_10 = mu_11$
*  $H_1: !H_0$. At least two are different.

```{r}
casesanova <- aov(Cases...Total~Month, data = dfnew) 
summary(casesanova)
```

 The F statistic is 35.25 and the p-value is 2e-16.
 P-value is small. We can conclude that there is a difference in the means of 
 the 9 groups. The result is that there is significant evidence that the means are not all equal.  

```{r}
TukeyHSD(casesanova, conf.level = .95)
plot(TukeyHSD(casesanova, conf.level = .95))
```

```{r}
boxplot(dfnew$Deaths...Total~dfnew$Month)
```

 looks like there might be some differences

*  $H_0: mu_3 = mu_4 = mu_5 = mu_6 = mu_7 = mu_8 = mu_9 = mu_10 = mu_11$
*  $H_1: !H_0$. At least two are different.

```{r}
deathsanova <- aov(Deaths...Total~Month, data = dfnew) 
summary(deathsanova)
```

```
The F statistic is 148.6 and the p-value is 2e-16.
```


P-value is small. We can conclude that there is a difference in the means of 
the 9 groups. The result is that there is significant evidence that the means are not all equal.  

```{r}
TukeyHSD(deathsanova, conf.level = .95)
plot(TukeyHSD(deathsanova, conf.level = .95))

```











